---
title: "Exploratory Analysis"
author: "Tianyou Wang"
date: "2022-10-11"
output: html_document
---

## *Setting Opotions*

From last lecture, in addition to figure sizing, we include a few other figure preferences in global options declared at the outset of each .Rmd file (this code chunk just gets copy-and-pasted to the beginning of every new file).

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```





## Load the Weather Data
`lubridate::floor_date(date, unit = "month")` round the days to the first day of the month.
```{r warning = FALSE}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```





##  *`group_by`*

Datasets are often comprised of groups defined by one or more (categorical) variable; group_by() makes these groupings explicit so that they can be included in subsequent operations. For example, we might group weather_df by name, which give us 3 groups.
```{r}
weather_df %>%
  group_by(name)
```


For example, we might group weather_df by name and month, which give us 36 groups.
```{r}
weather_df %>%
  group_by(name, month)
```


We can also ungroup after we group.
```{r}
weather_df %>%
  group_by(name, month) %>% 
  ungroup(month)
```
