---
title: "Exploratory Analysis"
author: "Tianyou Wang"
date: "2022-10-11"
output: html_document
---

## *Setting Opotions*

From last lecture, in addition to figure sizing, we include a few other figure preferences in global options declared at the outset of each .Rmd file (this code chunk just gets copy-and-pasted to the beginning of every new file).

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```





## Load the Weather Data
`lubridate::floor_date(date, unit = "month")` round the days to the first day of the month.
```{r warning = FALSE}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```





##  *`group_by`*

Datasets are often comprised of groups defined by one or more (categorical) variable; group_by() makes these groupings explicit so that they can be included in subsequent operations. For example, we might group weather_df by name, which give us 3 groups.
```{r}
weather_df %>%
  group_by(name)
```


For example, we might group weather_df by name and month, which give us 36 groups.
```{r}
weather_df %>%
  group_by(name, month)
```


We can also ungroup after we group.
```{r}
weather_df %>%
  group_by(name, month) %>% 
  ungroup(month)
```




## *Counting Things*

### `summarize()` 

We will frequently use `summarize()` to create one-number-variable that summaries within each group, or use `mutate()` to define variables within groups. 

For example, counting the number of observations in each month.
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n())
```


*** `count()`
We can also use `count()` in place of `group_by()` and `summarize()`.
```{r}
weather_df %>%
  count(name, month, name = "n_obs")
```


**Never** use base R's `table()` function!!!
`count()` is a useful tidyverse alternative to Base R’s `table()` function. Both functions produce summaries of how often values appear (number of occurrences), but table’s output is a classic n*2 table and is hard to do any additional work with it. `count()` adds one or more numeric columns to the original data frame, which can be used or manipulated later directly. For an example, run the code below and check the classic 12-by-2 table.

```{r}
weather_df %>%
  pull(month) %>% 
  table()
```



### Other helpful counters

We can use `summarize()` to compute *multiple summaries* within each group; creating multiple numeric columns. As an example, we count the number of observations in each month and the number of distinct dates in each month.
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```




## A digression on 2-by-2 tablee

Let’s say we want to look at the number of cold and not-cold days in Central Park and Waterhole. We can do this with some extra data manipulation steps through `group_by` and `summarize`.
Note: 

- `!=` means not equal to..
- `TRUE      ~ ""` if there is thing don't fill in any group, just assign ""

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```


We get a “tidy” table in the previous example, and that's also a data frame. We could re-organize it into a more standard (non-tidy) 2x2 table using `pivot_wider`, or you could use `janitor::tabyl(1-, 2-, or 3-way )`.
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  janitor::tabyl(name, cold)
```




## *General Summaries*

Standard statistical summaries are regularly computed in `summarize()` using functions like `mean()`, `median()`, `var()`, `sd()`, `mad()`, `IQR()`, `min()`, and `max()`. To use these, you put the variable to which they apply in the function. We can also include any additional arguments as necessary, such as `na.rm = TRUE` which get ride off the missing value.

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE))
```


If you want to summarize multiple columns using the same summary, the `across()` function is helpful. Under this example, R calculate mean for variable tmin to prcp.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```



The fact that `summarize()` produces a data frame is important. We can incorporate grouping and summarizing within broader analysis pipelines. For example, we can create a plot based on the monthly summary:
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE),
            mean_prec = mean(prcp, na.rm = TRUE),
            median_tmax = median(tmax, na.rm = TRUE)
            ) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() + 
  geom_line() + 
  theme(legend.position = "bottom")
```


Reminder: sometimes your results are easier to read in another format.
Note: `knitr::kable` is a table generator, which create tables in R_Markdown, Html, LaTex..... We can also add design decisions, such as `digits = 1` which saves numbers in 1 decimal place.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```







