---
title: "Visualization"
subtitle: "Part I"
author: "Tianyou Wang"
date: "2022-10-11"
output: html_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the Weather Data

```{r warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

What the r chunk did:

- searching online for the data we want
  - `rnoaa` package allows us to read public data online instead of reading excel, csv files
  - `meteo_pull_monitors(c())` pull the monitors we want, in this case we are getting weather data from "USW", "USC", and "USS" stations
  - `var = c()` bring the variables we want
  - `data_min` and `data_mx` are used to set the starting and ending date we are interested
- data manipulation through `mutate()`
  - `name = recode` is used to rename variables
  - `select()` is similiar to relocate, it's used to manipulate the column order





## Scatter Plots

```{r warning = FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

Typically, when we use `ggplot`

1. choose the data frame
2. `aes()`defining aesthetic mappings
3. choose the geomatrics



#### New approach, same graph.

- The code below could be used instead to produce the same figure. Using this style can be helpful if you want to do some pre-processing before making your plot but don’t want to save the intermediate data.
```{r warning = FALSE}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```



#### Also

We can also save the output of `ggplot()` to an object that can be printed and modified later.
```{r warning = FALSE}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()

```





## Advanced Scatterplot

#### `color` and `geom_smooth`

```{r warning = FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

How did we make it fancy:
- `color = cat_var` color by catagetorical variables
- `geom_smooth` additional geometry that can add to the plot which generate a smooth trajectery for the data


```{r warning = FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + 
  geom_smooth()
```
In this r chunk, colors only apply for the geom_point. Color is not defined in the general aesthetic mappings, so we only generated one smooth line through `geom_smooth`.



#### `alpha` in `geom_print`

```{r warning = FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

What we added:

- `facet_grid(. ~ col_var)` the `.` means the row variable is not defined here. The column variable is defined (~) as "name".
- `alpha = 0.5` in `geom_point` changes the transperancy level



#### `alpha` in global/general setting

```{r warning = FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax, alpha = tmin, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```
What we changed:

- `alpha = tmin` in `aes()` gives a color gradient in the plot. the smaller "tmin" has more transparent color



#### New plot: Sizing the data point based on a variable

We sized the data point based on the "prcp" variable, where the higher "prcp" the bigger the data point showed on graph, via `geom_point(aes(size = prcp))`.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~name)
```





## Some Notes

### We can have whatever geoms we want. The number of geoms is not limited.

##### Having one gemo.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
```



##### Having two gemos.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```



### We can have a neat geom. 

##### `geom_hex()`
The `geom_hex()` shows as hexagens.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()
```



##### Or `geom_bin2d()`
The `geom_bin2d()` shows as squares.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_bin2d()
```



##### Or something looks like a elevation map
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_density2d() + 
  geom_point(alpha = 0.3)
```





## Univariate Plots

### Histograms

```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Add color in histogram.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram()
```
This graph looks weired, so we want to change:

- fill the color by using `fill = name` instead of `color = name`
- change the histograms' positions and make them aside to each other, via `position = "dodge"`


```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```


We can also change the position by separating the graph into 3 graphs, where each graph contains histograms for one location only.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```



### New geometry `geom_density`

`geom_density` kind of smooth out the edges of the histograms. It shows the overall shape of the distribution, but we will lose those little bumps.
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.4)
```



### Boxplots

```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(y = tmin)) +
  geom_boxplot()
```

We stratify the data points based on locations ("name").
```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```



### Trendy Plots

##### Violin Plot

```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "median")
```
Turn your head for 90 degrees, on the side, the violin plot show you the distribution.

We can add statistic features/functions in the plot by using `stat_summary(fun = "median")`. In this case we added one of the summary statistics functions, the median.


##### Ridge plot - the most popular plot of 2017

```{r warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges(alpha = 0.5)
```
Ridge plots show you the distributions. 





## Save and Embed

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```


What about embedding...
```{r, fig.width=6, fig.height=3}
weather_plot
```


Embed at a different size.
```{r, fig.width=12, fig.height=6}
weather_plot
```
