---
title: "Visualization"
subtitle: "Part II"
author: "Tianyou Wang"
date: "2022-10-11"
output: html_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the Weather Data

```{r warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```



## *Labels*

Providing informative axis labels, plot titles, and captions, via `labs(title = "", x = "", y = "", caption = "")`.
```{r Warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  )
```



## *Scales*
We can set scales for x and y axis by using `scale_x_continuous()`. We can also label them by using `labels = c()` inside of `scale_x_continuous()`.

```{r Warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15ºC")) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

What we did for the x, y-axis scales:

- `breaks = c()` setting the scale points, seperated by commas
- `labels = c()` choice how we label the scale points, such as adding units
- `trans = ""` is used to transform the data, such as trans = "sqrt" (square root) and trans = "log"
- `position = "right"` putting the scale discription on the right side of graph



### Color Scales
```{r Warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
  scale_color_hue(
    name = "Location",
    h = c(100, 300))
```

Using `name = "xxx"` to rename the coloring variance, in this case we use 
(name = "Location") to change our variable "name" to "Location". Note: the variable "name" accidentally shared the same name with the (name = "xxx") command.

The default hue is "h = c(1, 360)"



### Viridis Package for Color Scheme
Trying to create your own color scheme usually doesn’t go well. You can use the "viridis package" instead. There are several options, but the default color scheme works nicely!

Note: `viridis::scale_color_viridis()` assume the variable you put in is a continuous variable. If we are applying it to a discrete variable (this case, location,) we want to write `discrete = TRUE` inside the scale_color_viridis command.

```{r Warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```





## *Themes*
Themes are used to modify *non-data elements of a plot*. They don’t change mappings or how data are render, but control things like background color and location of the the legend. Using themes can help with general plot appearance.

For example, the default legent position is on the right of the graphic. If we want to move it to the bottom, we can use `theme(legend.position = "bottom")`.

```{r Warning = FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```



Or we can name the plot and add themes.
```{r Warning = FALSE}
ggp_temp_plot =
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```

```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom")
```



Some other theme examples:
White background with black box around.
```{r}
ggp_temp_plot + 
  theme_bw()
```


Just white background.
```{r}
ggp_temp_plot + 
  theme_minimal() +
  theme(legend.position = "bottom")

```


```{r}
ggp_temp_plot + 
  theme_classic()
```

In `ggthemes` there are a lot of themes.
```{r}
ggp_temp_plot + 
  ggthemes::theme_economist_white()
```


```{r}
ggp_temp_plot + 
  ggthemes::theme_excel()
```


```{r}
ggp_temp_plot + 
  ggthemes::theme_excel_new()
```





## *Setting Opotions*

In addition to figure sizing, I include a few other figure preferences in global options declared at the outset of each .Rmd file (this code chunk just gets copy-and-pasted to the beginning of every new file).

```{r eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

What we did in this code chunk:

- `opt_chunk$set()` setting the R code chunks
- `fig.width` sets the figure width
- `fig.asp` set the aspect ratio (height/width)
- `out.width` let the figure sit bit away from the margins
- `theme_set` setting the graphic themes
- setting color and fill preferences globally through `options()`





## *Data Argument in `geom`*

It’s sometimes necessary to overlay data summaries on a plot of the complete data. Depending on the setting, one way to do this is to create a “summary” dataframe and use that when adding a new geom to a ggplot based on the full data.


We can split "weather_df" into separate datasets for Central Park and Waikiki. Then we use one dataset in the `ggplot()` and use another in `geom_line()` inside of `ggplot()`. 


Note: we only changed the data set used in `geom_line()`, it still adopt the global `aes()` settings.

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```





## *`patchwork`*

Remember faceting? We’ve seen facetting as an approach to create the “same type of plots” for several levels of a categorical variable.

```{r warning=FALSE}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~name)
```

What happen when you want multiple plots but cannot facet...?
Sometimes, we want to combine two or three fundamentally different plots in the same graphic, such as showing a scatterplot and a boxplot, or show scatterplots illustrating relationships between different variables. In this case, the solution is to create each of the panels you want separately and combine panels using tools in the `patchwork` package which is loaded in the beginning of this R_Markdown file.

```{r warning=FALSE}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```





## *Data Manipulation*

The behavior of plot depends on the data we have supplied. In some cases, it’s easier to control behavior through data manipulation than it is through the plot code.

This is particularly true for the order of categorical or factor variables in plots (this case, "name" variable.) Categorical variables will be ordered alphabetically. Factors will follow the specified order level that underlies the variable labels. We can change the order level of a factor variable to your specified preference using `forcats::fct_relevel` or according to the value of another variable using `forcats::fct_reorder`.

```{r warning=FALSE}
weather_df %>%
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```


Note: changing the categorical "name" to factor, via `name = factor(name)` is not necessary.
```{r warning=FALSE}
weather_df %>%
  mutate(
    name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```


We reorders "name" according to tmax values in each "name."
```{r warning=FALSE}
weather_df %>%
  mutate(
    name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```




## *Other Examples*

### Weather_df

What if We wanted density graph for tmin and tmax simultaneously?

We want to facet panels across the "name" variable, and create separate density plots for tmax and tmin in each panel. Unfortunately, weather_df isn’t organized in a way that makes this easy.

One solution would recognize that tmax and tmin are separate observation types of a shared temperature variable. With this understanding, we can shift the data into long format which can be used to make the plot directly:

```{r warning=FALSE}
weather_df %>%
  select(name, tmax, tmin) %>%
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>%
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```


If we only want to make the graph for CentralPark_NY, we can fillter the info out from the weather_df first.

```{r warning=FALSE}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>%
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = 0.5)
```


### PULSE Data

The code below imports and tidies the PULSE data, and creates boxplots showing BDI score across visits. Using `pivot_longer` to organize the BDI score and visit time variables, and organizing the visit time variable into a factor with an informative ordering.

```{r warning=FALSE}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi, fill = visit)) + 
  geom_boxplot()
```

Note: `visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))` is used to order the factor level. So, we see "00m" at the most left of the graph, and "12m" locates at the most right of the graph.



### FAS Data

We’ve seen code for data import and organization and for joining the litters and pups data. Here we add some data tidying steps to view pup-level outcomes (such as, post-natal day on which ears “work”, on which the pup can walk, etc.) across values of dose category and treatment day.

```{r warning=FALSE}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```

Note:

1. `select(sex, dose, day_of_tx, pd_ears:pd_walk)` can be achieved by `select(sex, dose, day_of_tx, starts_with("pd_))` too.

2. `forcats::fct_reorder(outcome, pn_day, median)` let R reorder "outcome" based on the median "pn_day". It can also be achieved, through a less general way, via `forcats::fct_reorder(outcome, "pd_ears", "pd_pivot", "pd_walk", "pd_eyes")`.

