---
title: "Visualization"
subtitle: "Part I"
author: "Tianyou Wang"
date: "2022-10-11"
output: html_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the Weather Data

```{r warning = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

What the r chunk did:

- searching online for the data we want
  - `rnoaa` package allows us to read public data online instead of reading excel, csv files
  - `meteo_pull_monitors(c())` pull the monitors we want, in this case we are getting weather data from "USW", "USC", and "USS" stations
  - `var = c()` bring the variables we want
  - `data_min` and `data_mx` are used to set the starting and ending date we are interested
- data manipulation through `mutate()`
  - `name = recode` is used to rename variables
  - `select()` is similiar to relocate, it's used to manipulate the column order



## Scatter Plots

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

Typically, when we use `ggplot`

1. choose the data frame
2. `aes()`defining aesthetic mappings
3. choose the geomatrics

#### New approach, same graph.

- The code below could be used instead to produce the same figure. Using this style can be helpful if you want to do some pre-processing before making your plot but don’t want to save the intermediate data.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```


#### Also

We can also save the output of `ggplot()` to an object that can be printed and modified later.
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()

```

